---
title: "HPC"
author: "Jeremie Alexander"
date: "13/03/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load required libraries

```{r}
library(parallel)
library(foreach)
library(doParallel)
```


Run the detectCores() function from the parallel package

```{r}
detectCores()
```


Write a standard for loop 4,000 times, and each time calculate the mean of 100,000 random numbers from a normal distribution with mean of 10 and standard deviation of 3

```{r}

StartTime = Sys.time()
means = c()
for (i in 1:4000) {
  random_numbers <- rnorm(100000, mean = 10, sd = 3)
  means[i] <- mean(random_numbers)
}
EndTime = Sys.time()

#Obtain start/end times
StartTime
EndTime

TimeDiff = EndTime - StartTime
TimeDiff
```

Thus, the total run time was ~11.5 seconds

Predict how much faster a multi-threaded program will run compared to a serial program, based on the number of threads you have???

```{r}
PredictedTimeDecrease = TimeDiff/detectCores()
PredictedTimeDecrease
```

Based on this, I predict the the parallel run take 1.56 seconds

Now modify your for loop to run the same code in parallel, instead of running each of the 4,000 loops in order.

```{r}

Cores <- parallel::makeCluster(detectCores())
doParallel::registerDoParallel(Cores)

StartTimePar = Sys.time()
meansPar <- foreach(i = 1:4000, .combine = rbind) %dopar% {
  random_numbers2 <- rnorm(100000, mean = 10, sd = 3)
  mean(random_numbers2)
}
EndTimePar = Sys.time()
parallel::stopCluster(Cores)

#Obtain start/end times
StartTimePar
EndTimePar

TimeDiffPar = EndTimePar - StartTimePar
TimeDiffPar
```

Parallel start time: "2023-03-13 16:00:53 EDT"
Parallel end time: "2023-03-13 16:00:58 EDT"

Thus, the parallel run time was ~4 seconds

Compare the run-time for your parallel loop with the run-time for your serial loop and the theoretical run-time you calculated for the parallel loop based on the number of cores in your computer

```{r}

```



















